\textls[-15]{\thetitle \mbox{(Technical Report)}}

### 2022-11-09 casestudies.tex
edits on AI case study

The code is organized into four families.
A base family, \lsti{Imp}, defines the abstract syntax of a while-language
with pure expressions and impure statements, via \lsti{FInductive}.
The semantic is formulated as a functional interpreter like a CEK machine\cite{felleisen1986control}, parameterized by a fuel value,
via \lsti{FRecursion}.
It contains \textasciitilde 200~LOC.

A second family, \lsti{ImpGAI}, extends \lsti{Imp} and contains
\textasciitilde 550~LOC.
It exports a generic framework for deriving
abstract interpreters with partial correctness.
The soundness theorem of this generic abstract interpreter, \lsti{analyze},
is with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
It states that the abstraction relation \lsti{RState} over a
concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
by the analysis:

\begin{centered}
\begin{minipage}{.86\textwidth}
\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze fuel stmt absS)
\end{lstlisting}
\end{minipage}
\end{centered}

@@

\noindent
\lsti{analyze} is defined via \lsti{FRecursion}, and the soundness theorem
is proved via \lsti{FInduction}.
This family leaves fields representing the abstract domain, the
abstraction relation, monotonicity of transfer functions, etc.\ 
largely unspecified or unproven\EDJ{It is not very relaxed. Currently everything is using a dicionary as memory model and stick with that. The abstract value for each expression is the one *really* unspecified.}---a derived family can instantiate these
``parameters'' by overriding appropriate fields (and also possibly extend the abstract syntax),
to create a sound, runnable abstract interpreter for a (possibly extended) while-language.

The other two families then extend \lsti{ImpGAI}.
Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
interpreter that does type inference~\cite{cousot1997types}.
Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
with natural number arithmetic,
and instantiates the generic abstract interpreter to perform constant propagation.

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
directly ready for program extraction to OCaml and put them to test over queries.

We note that our plugin cannot yet allow the expression of two abstract
interpreters nested within the same family.
Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
to further increase the expressive power of our language design.

>>>>

The code is organized into four families.
A base family, \lsti{Imp}, defines via \lsti{FInductive} the abstract
syntax of a while-language with pure expressions and impure statements.
The semantics is given by an interpreter defined as a CEK-style abstract
machine~\cite{felleisen1986control} and parameterized by a fuel value.
Family \lsti{Imp} defines the interpreter via \lsti{FRecursion}.
It contains \textasciitilde 200~LOC.

A second family, \lsti{ImpGAI}, extends \lsti{Imp} and uses
\textasciitilde 550~LOC.
It exports a generic framework for deriving
abstract interpreters with partial-correctness guarantees.
Soundness of the abstract interpreter, \lsti{analyze},
is stated with respect to the interpreter, \lsti{eval}, inherited from \lsti{Imp}.
The theorem says that the abstraction relation \lsti{RState} over a
concrete state~\lsti{S} and an abstract state~\lsti{absS} is preserved
by the analysis:

\begin{centered}
\begin{minipage}{.91\textwidth}
\begin{lstlisting}[basicstyle=\fontsize{8.25}{9}\ttfamily]
\forall stmt fuel S absS, RState S absS -> RState (eval fuel stmt S) (analyze fuel stmt absS)
\end{lstlisting}
\end{minipage}
\end{centered}

@@

\noindent
\lsti{analyze} is defined via \lsti{FRecursion}, and the soundness theorem
is proved via \lsti{FInduction}.
This family leaves fields representing the abstract domain, the
abstraction relation, monotonicity of transfer functions, etc.\ 
largely unspecified or unproven---a derived family can further bind
these ``parameters'' by overriding appropriate fields (and also possibly
extend the abstract syntax), to create a sound, runnable abstract
interpreter for a (possibly extended) while-language.
%\EDJ{It is not very relaxed. Currently everything is using a dicionary as memory model and stick with that. The abstract value for each expression is the one *really* unspecified.}

The other two families then extend \lsti{ImpGAI}.
Family \lsti{ImpTI} (\textasciitilde200~LOC) is an abstract
interpreter that does type inference~\cite{cousot1997types}.
Family \lsti{ImpCP} (\textasciitilde300~LOC) extends the abstract syntax
with natural number arithmetic,
and further binds the generic abstract interpreter to perform constant propagation.

Our implementation of family polymorphism retains the computation
ability of the host proof assistant;
the two abstract interpreters in \lsti{ImpTI} and \lsti{ImpCP} are
ready for program extraction to OCaml.
We put them to test over queries.

We note that our plugin cannot yet allow the expression of two derived abstract
interpreters nested within the same family.
Work on nested inheritance~\cite{ncm2004,zm2017} points to a direction
to further increase the expressive power of our language design.

### 2022-11-09 coqimpl.tex
edits on AI case study

Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a module),
as the last line of \cref{fig:stlcfix-compiled} shows.

>>>>

Finally, the reference \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a family)
can simply be translated to \lsti{STLCFix.typesafe} (where \lsti{STLCFix} is a Coq module),
as the last line of \cref{fig:stlcfix-compiled} shows.
